---
title: Independent Study - Spatial Statistics
author: "Prakkash Manohar, IIIT Sri City"
date: "Oct 15, 2018"
output:
  html_document:
    df_print: paged
  html_notebook: default
  pdf_document: default
---
**<u><font size=5>LOADING THE DATASET</font></u>**
```{r}
#  D I W A L I    D A T A S E T  - - - - - - - - - - - - - - - - - - - - - - - - -
#setwd("C:/Users/lenovo/Desktop/Honors Evaluation/Work")
rm(list = ls())
options(scipen = 999)
library(readxl)
library(geoR)
library(ggplot2)
library(pracma)
library(RandomFields)

source('C:/Users/lenovo/Desktop/Honors Evaluation/Work/vario.r')
source('C:/Users/lenovo/Desktop/Honors Evaluation/Work/varifit.r')

data3 <- read_excel("data2.xlsx")
# X11()
# plot(data3$X, data3$Y)
# # set.seed(123)
# # tmpindex <- sample(1:nrow(data3), .3*nrow(data3))
# tmpindex <- identify(data3$X, data3$Y)
tmpindex <- c(4,9,13,14,18,24,27,30)
predict_data <- data3[tmpindex,]
data3 <- data3[-tmpindex,]
a <- as.vector(data3[,2])
b <- as.vector(data3[,3])
c <- as.vector(data3[,4])
rm(data3)

data3 <- data.frame(a, b, c)
# data2[] <- lapply(data2, function(x) {
#     if(is.factor(x)) as.numeric(as.character(x)) else x
# })
geodata1 <- as.geodata(data3, coords.col = 1:2, data.col = 3)


# my function
ideal_variogram_2 <- variogram(data3[,1], data3[,2], data3[,3], 18, 0, 0.025, 90, 45, 1000000)
ideal_variogram_2 <-  na.omit(ideal_variogram_2)
plot(ideal_variogram_2$lag, ideal_variogram_2$gamma, main = "Using my code")

ideal_variogram_1 <- variog(geodata1, coords = geodata1$coords, data = geodata1$data, option = "bin", uvec = seq(0, 0.45, by = 0.025), direction = 0, tolerance = pi/4)
df <- data.frame(u = ideal_variogram_1$u, v = ideal_variogram_1$v)
plot(df$u, df$v, main = "Using variog()", ylim = c(0, 4000))

library(xlsx)
write.xlsx(df, "df.xlsx")

# for(i in seq(0.001, 0.04, by = 0.001)){
#   ideal_variogram_1 <- variog(geodata1, coords = geodata1$coords, data = geodata1$data, option = "bin", uvec = seq(0, 0.45, by = i), direction = 0, tolerance = pi/4)
#   df <- data.frame(u = ideal_variogram_1$u, v = ideal_variogram_1$v)
# 
# plot(df$u, df$v, main = i, ylim = c(0, 4000))
# }

# choosing spherical variogram parameters by ordinary least squares (OLS)

# min <- Inf
# for(c0 in seq(0, by = 0.001, 0.06)){
#   for(sill in seq(0.05, by = 0.001, 0.1)){
#     for(a in seq(50, by = 5, 300)){
#       idealvariogram <- varifit(c0, sill, a, 1, ideal_variogram_2$lag)
#       tmp <- (idealvariogram - ideal_variogram_2$gamma)^2
#       rmse <- sqrt(mean(na.omit(tmp[1:117])))
#       if(rmse < min)
#         min <- rmse
#         bestpar <- c(c0, sill, a)
#       }
#     }
#   }

predict_data <- predict_data[,c(2:4)]
a <- as.vector(predict_data[,1])
b <- as.vector(predict_data[,2])
c <- as.vector(predict_data[,3])
rm(predict_data)
predict_data <- data.frame(a, b, c)

predicted_data <- cbind(predict_data, "Predicted_PM2.5_Kriging" = 0)
dist <- pairdist(data3$X, data3$Y, na.ignore = T)

bestpar <- c(300, 1500, 0.25)        # found from the matlab program

c0 <- bestpar[1]
c1 <- bestpar[2] - c0
a <- bestpar[3]

tmp <- varifit(c0, c0+c1, a, 1, df$u)
fit <- data.frame("u" = df$u, "v" = tmp)

ggplot(df, aes(x = u, y = v)) +
  geom_point(shape = 3, size = 2, colour = "black") +
  ggtitle("Variogram for PM2.5 data") +
  labs(x = "Distance", y = "Semivariance/Moment of inertia")

ggplot() +
  geom_point(data = df, mapping = aes(x = u, y = v), col = "black") +
  geom_line(data = fit, mapping = aes(x = u, y = v), shape = 16, colour = "red", linetype="solid") +
  ggtitle("Fitted Spherical Variogram") +
  labs(x = "Distance", y = "Semivariance/Moment of inertia")



# # PLOTTING THE VARIOGRAM FOR DIFFERENT DATES (04-11-2018 to 11-11-2018)

# data4 <- predictData
# for(i in 3:10){
#   tmp <- na.omit(data4[,c(1,2,3)])
#   geodata1 <- as.geodata(tmp, coords.col = 1:2, data.col = 3)
#   geodata1
#   v1 <- variog(geodata1, coords = geodata1$coords, data = geodata1$data, option = "bin", uvec = seq(0, 0.45, by = 0.025), direction = pi/2, tolerance = pi/4)  #, breaks = seq(0, 1400, l = 21)
#   plot(v1)
#   plot(v1$u, v1$v)
#   #a <- cat("plot", i, ".png")
#   tmp <- as.numeric(unlist(na.omit(data4[,i])))
#   plot(density(tmp))
#   qqnorm(tmp)
#   qqline(tmp, col = "green")
#   #plot(v1$u, v1$v, main = "Variogram for Delhi PM2.5 Data", xlab = "Distance", ylab = "Semivariance / Moment of inertia", pch = 20)
#   #dev.copy(png,a)
#   #dev.off()
# }

# predicted_data$Predicted_PM2.5 <- kriging(as.matrix(data3[,c(1:2)]), data3[,3], as.matrix(predicted_data[,c(1:2)]), type = "ordinary")
# my_RMSE<-caret::RMSE(predicted_data$Predicted_PM2.5, predicted_data$PM2.5)

#predicted_data$Predicted_PM2.5 <- Kriging(krige.method = "O", x = data3[,1], y = data[,2], model = "spherical")

# KRIGING (my code)

covariance1 <- ifelse(dist > a, 0, c1*(1 -(1.5*abs(dist)/a)+(0.5*(abs(dist)/a)^3)))

covariance1 <- cbind(covariance1, rep(1, nrow(covariance1)))
covariance1 <- rbind(covariance1, rep(1, ncol(covariance1)))
covariance1[nrow(covariance1), ncol(covariance1)] <- 0

covinverse <- inv(covariance1)       #finds inverse of the matrix

dist2 <- proxy::dist(predict_data[1:2], data3[1:2])

covariance2 <- ifelse(dist2 > a, 0, c1*(1 -(1.5*abs(dist2)/a)+(0.5*(abs(dist2)/a)^3)))
# covariance2 <- rbind(covariance2, 1)

data4 <- t(rbind(data3[3], 0))
d <- t(as.matrix(covariance2))
d <- rbind(d, 1)
w <- covinverse %*% d
predicted_data["Predicted_PM2.5_Kriging"] <- as.vector(data4 %*% w)
predicted_data["Predicted_PM2.5_Kriging"] <- sapply(predicted_data["Predicted_PM2.5_Kriging"], function(x) ifelse(x < 0, 0, round(x, 2)))
#print(w)

```

```{r}
library(phylin)
#library(spatstat)
#library(gstat)

# PREDICTING FOR THE ENTIRE GRID USING KRIGING

# x.range <- as.numeric(c(round(min(data3$X), 2), round(max(data3$X), 2)))  # min/max longitude of the interpolation area
# y.range <- as.numeric(c(round(min(data3$Y), 2), round(max(data3$Y), 2)))  # min/max latitude of the interpolation area

# GRID
# from the range, exapnd the coordinates to make a regular grid
# grd <- expand.grid(x = seq(from = x.range[1], to = x.range[2], by = 0.001), y = seq(from = y.range[1], to = y.range[2], by = 0.001))
library(raster)
library(rgdal)
library(rgeos)
library(ggplot2)
library(dplyr)

### Q1: Delhi only

### Get data
india <- getData("GADM", country = "India", level = 1)

### Choose Delhi
delhi <- subset(india, NAME_1 == "NCT of Delhi")

### SPDF to DF
delhiline <- fortify(delhi)

grid <- makegrid(delhi, cellsize = 0.001) # cellsize in map units!
grid <- SpatialPoints(grid, proj4string = CRS(proj4string(delhi)))

grid <- grid[delhi, ]
grid <- as.data.frame(grid)

# KRIGING

covariance1 <- ifelse(dist > a, 0, c1*(1 -(1.5*abs(dist)/a)+(0.5*(abs(dist)/a)^3)))

covariance1 <- cbind(covariance1, rep(1, nrow(covariance1)))
covariance1 <- rbind(covariance1, rep(1, ncol(covariance1)))
covariance1[nrow(covariance1), ncol(covariance1)] <- 0

covinverse <- inv(covariance1)       #finds inverse of the matrix

dist2 <- proxy::dist(grid[1:2], data3[1:2])

covariance2 <- ifelse(dist2 > a, 0, c1*(1 -(1.5*abs(dist2)/a)+(0.5*(abs(dist2)/a)^3)))
# covariance2 <- rbind(covariance2, 1)

data4 <- t(rbind(data3[3], 0))
d <- t(as.matrix(covariance2))
d <- rbind(d, 1)
w <- covinverse %*% d
predicted_data_2 <- as.vector(data4 %*% w)
predicted_data_2 <- sapply(predicted_data_2, function(x) ifelse(x < 0, 0, round(x, 2)))
#print(w)

kriging.output <- data.frame(as.vector(grid$x1), as.vector(grid$x2), predicted_data_2)
colnames(kriging.output) <- c("grd.x", "grd.y", "PM2.5")

# IDW

predicted_data["Predicted_PM2.5_IDW"] <- round(phylin::idw(data3[,3], data3[,c(1,2)], predict_data[,c(1,2)]), 2)

interpolated <- phylin::idw(data3[,3], data3[,c(1,2)], grid)
idw.output <- data.frame(as.vector(grid$x1), as.vector(grid$x2), as.vector(interpolated))
colnames(idw.output) <- c("grd.x", "grd.y", "PM2.5")

ggplot() + 
  geom_tile(data = kriging.output, aes(x = grd.x, y = grd.y, fill = PM2.5)) + 
  geom_point(data = data3, aes(x = X, y = Y), shape = 19, color = "black") +
  geom_point(data = predict_data, aes(x = X, y = Y), shape = 3, color = "black") +
  scale_fill_distiller(palette = "Paired", direction = 1) +
  theme_bw() +
  
  labs(title = "Kriging")

ggplot() + 
  geom_tile(data = idw.output, aes(x = grd.x, y = grd.y, fill = PM2.5)) + 
  geom_point(data = data3, aes(x = X, y = Y), shape = 19, color = "black") +
  geom_point(data = predict_data, aes(x = X, y = Y), shape = 3, color = "black") +
  scale_fill_distiller(palette = "Paired", direction = 1) +
  theme_bw() +
  
  labs(title = "IDW")

# Palettes
# Diverging - BrBG, PiYG, PRGn, PuOr, RdBu, RdGy, RdYlBu, RdYlGn, Spectral
# Qualitative - Accent, Dark2, Paired, Pastel1, Pastel2, Set1, Set2, Set3
# Sequential - Blues, BuGn, BuPu, GnBu, Greens, Greys, Oranges, OrRd, PuBu, PuBuGn, PuRd, Purples, RdPu, Reds, YlGn, YlGnBu, YlOrBr, YlOrRd
```

<font size = 5>**There are many things which can be observed/inferred from the above variogram, in general:**<br>
<ol><li> The moment of inertia/semivariance is less for low distance, which is understandable as there will be many data points close to each other (for low distances).

<li> Theoretically, as distance increases, data points drift away from the line _x=y_ on an h-scatterplot so the cloud of data points will get fatter and MOI increases. This can be verified from this variogram. As the distance increases, the MOI also increases (roughly).</ol></font>

```{r}
# ERROR METRICS
library(ModelMetrics)
library(MLmetrics)
mae1 <- mae(predicted_data$PM2.5, predicted_data$Predicted_PM2.5_Kriging)
mae2 <- mae(predicted_data$PM2.5, predicted_data$Predicted_PM2.5_IDW)

my_RMSE_1 <- caret::RMSE(predicted_data$Predicted_PM2.5_Kriging, predicted_data$PM2.5)
my_RMSE_2 <- caret::RMSE(predicted_data$Predicted_PM2.5_IDW, predicted_data$PM2.5)

mape1 <- MAPE(predicted_data$Predicted_PM2.5_Kriging, predicted_data$PM2.5)
mape2 <- MAPE(predicted_data$Predicted_PM2.5_IDW, predicted_data$PM2.5)

df1 <- data.frame("Ordinary Kriging" = c(mae1, my_RMSE_1, mape1), "IDW Interpolation" = c(mae2, my_RMSE_2, mape2))
colnames(df1) <- c("Ordinary Kriging", "IDW Interpolation")
rownames(df1) <- c("Mean Absolute Error (MAE)", "Root Mean Square Error (RMSE)", "Mean Absolute Percentage Error (MAPE)")

df1 <- round(df1, 2)

df1[3,] <- lapply(df1[3,], function(x) paste(as.numeric(x)*100, "%", sep = ""))

```

```{r echo = FALSE, warning = FALSE}
library(knitr)
library(kableExtra)
options(knitr.table.format = "html")
kable(df1, align="c", "html") %>%
kable_styling("striped",full_width = F) %>%
row_spec(1, color = "black", background = "#ffe6e6") %>%
row_spec(2, color = "black", background = "#ffff99") %>%
row_spec(3, color = "black", background = "#ffe6e6") %>%
column_spec(1, bold = T, border_right = T) %>%
column_spec(2, border_right = T)

```
