---
title: Independent Study - Spatial Statistics
author: "Prakkash Manohar, IIIT Sri City"
date: "Oct 15, 2018"
output:
  html_document:
    df_print: paged
  html_notebook: default
  pdf_document: default
---
**<u><font size=5>LOADING THE DATASET</font></u>**
```{r}
#  D I W A L I    D A T A S E T  - - - - - - - - - - - - - - - - - - - - - - - - -
setwd("F:/Logs/S_6 (S-2019)/Honors/Evaluation-1")
rm(list = ls())
options(scipen = 999)
library(readxl)
library(geoR)
library(ggplot2)
library(pracma)
library(RandomFields)
library(xlsx)

source('F:/Logs/S_6 (S-2019)/Honors/Evaluation-1/vario.r')
source('F:/Logs/S_6 (S-2019)/Honors/Evaluation-1/varifit.r')

data3 <- read_excel("data3.xlsx")
# X11()
# plot(data3$X, data3$Y)
# # set.seed(123)
# # tmpindex <- sample(1:nrow(data3), .3*nrow(data3))
# tmpindex <- identify(data3$X, data3$Y)
data3 <- cbind(data3[2], data3[1], data3[3:ncol(data3)])
tmpindex <- c(4,9,13,14,18,24,27,30)
predict_data <- data3[tmpindex,]
data3 <- data3[-tmpindex,]

i <- 0
hvalues <- c(0, 0, 0.013, 0.0067, 0.028, 0.02, 0.026, 0.019, 0.023, 0.016)

for(i in 3:ncol(data3)){
  data <- data.frame(data3[1], data3[2], data3[i])
  geodata1 <- as.geodata(data, coords.col = 1:2, data.col = 3)
  ideal_variogram_1 <- variog(geodata1, coords = geodata1$coords, data = geodata1$data, option = "bin", uvec = seq(0, 0.45, by = hvalues[i]), direction = 0, tolerance = pi/4)
  df <- data.frame(u = ideal_variogram_1$u, v = ideal_variogram_1$v)
  plot(df$u, df$v, main = "Using variog()")
  abline(h = var(na.omit(data[,3])), col="blue")
  #filename <- paste("df", i, ".xlsx", sep = "")
  #write.xlsx(df, filename)
}

# for(i in seq(0.001, 0.04, by = 0.001)){           # seq(0.001, 0.04, by = 0.001)               0.01, by = 0.0001 
#   ideal_variogram_1 <- variog(geodata1, coords = geodata1$coords, data = geodata1$data, option = "bin", uvec = seq(0, 0.45, by = i), direction = 0, tolerance = pi/4)
#   df <- data.frame(u = ideal_variogram_1$u, v = ideal_variogram_1$v)
# 
#   plot(df$u, df$v, main = i)
#   abline(h = var(data[,3]), col="blue")
# }
  
```

```{r}

# choosing spherical variogram parameters by ordinary least squares (OLS)

# min <- Inf
# for(c0 in seq(0, by = 0.001, 0.06)){
#   for(sill in seq(0.05, by = 0.001, 0.1)){
#     for(a in seq(50, by = 5, 300)){
#       idealvariogram <- varifit(c0, sill, a, 1, ideal_variogram_2$lag)
#       tmp <- (idealvariogram - ideal_variogram_2$gamma)^2
#       rmse <- sqrt(mean(na.omit(tmp[1:117])))
#       if(rmse < min)
#         min <- rmse
#         bestpar <- c(c0, sill, a)
#       }
#     }
#   }

# 3:    [650;1050;0.150000000000000]
# 4:    [1000;7100;0.100000000000000]
# 5:    [250;1350;0.208000000000000]
# 6:    [1300;3100;0.141000000000000]
# 7:    [5500;7700;0.200000000000000]
# 8:    [1000;3300;0.173000000000000]
# 9:    [788.37;4200;0.224000000000000]
# 10:   [2550;8400;0.075000000000000]

dist <- pairdist(data3$X, data3$Y, na.ignore = T)

j <- 3    # COLUMN NUMBER
predict_data_tmp <- data.frame(predict_data[, 1], predict_data[, 2], predict_data[, j])

bestpar <- c(650, 1050, 0.15)        # found from the matlab program

c0 <- bestpar[1]
c1 <- bestpar[2] - c0
a <- bestpar[3]

data <- data.frame(data3[1], data3[2], data3[j])
geodata1 <- as.geodata(data, coords.col = 1:2, data.col = 3)
ideal_variogram_1 <- variog(geodata1, coords = geodata1$coords, data = geodata1$data, option = "bin", uvec = seq(0, 0.45, by = hvalues[j]), direction = 0, tolerance = pi/4)
df <- data.frame(u = ideal_variogram_1$u, v = ideal_variogram_1$v)

tmp <- varifit(c0, c0+c1, a, 1, df$u)
fit <- data.frame("u" = df$u, "v" = tmp)

ggplot(df, aes(x = u, y = v)) +
  geom_point(shape = 3, size = 2, colour = "black") +
  ggtitle("Variogram for PM2.5 data") +
  labs(x = "Distance", y = "Semivariance/Moment of inertia")

ggplot() +
  geom_point(data = df, mapping = aes(x = u, y = v), col = "black") +
  geom_line(data = fit, mapping = aes(x = u, y = v), shape = 16, colour = "red", linetype="solid") +
  ggtitle("Fitted Spherical Variogram") +
  labs(x = "Distance", y = "Semivariance/Moment of inertia")

# predicted_data$Predicted_PM2.5 <- kriging(as.matrix(data3[,c(1:2)]), data3[,3], as.matrix(predicted_data[,c(1:2)]), type = "ordinary")
# my_RMSE<-caret::RMSE(predicted_data$Predicted_PM2.5, predicted_data$PM2.5)

#predicted_data$Predicted_PM2.5 <- Kriging(krige.method = "O", x = data3[,1], y = data[,2], model = "spherical")

# # KRIGING (for calculating error metrics)
# 
# covariance1 <- ifelse(dist > a, 0, c1*(1 -(1.5*abs(dist)/a)+(0.5*(abs(dist)/a)^3)))
# 
# covariance1 <- cbind(covariance1, rep(1, nrow(covariance1)))
# covariance1 <- rbind(covariance1, rep(1, ncol(covariance1)))
# covariance1[nrow(covariance1), ncol(covariance1)] <- 0
# 
# covinverse <- inv(covariance1)       #finds inverse of the matrix
# 
# dist2 <- proxy::dist(predict_data[1:2], data3[1:2])
# 
# covariance2 <- ifelse(dist2 > a, 0, c1*(1 -(1.5*abs(dist2)/a)+(0.5*(abs(dist2)/a)^3)))
# # covariance2 <- rbind(covariance2, 1)
# 
# data4 <- t(rbind(data3[3], 0))
# d <- t(as.matrix(covariance2))
# d <- rbind(d, 1)
# w <- covinverse %*% d
# predicted_data["Predicted_PM2.5_Kriging"] <- as.vector(data4 %*% w)
# predicted_data["Predicted_PM2.5_Kriging"] <- sapply(predicted_data["Predicted_PM2.5_Kriging"], function(x) ifelse(x < 0, 0, round(x, 2)))
# #print(w)


```

```{r}
library(phylin)
#library(spatstat)
#library(gstat)

# PREDICTING FOR THE ENTIRE GRID USING KRIGING

x.range <- as.numeric(c(round(min(data3$X), 2), round(max(data3$X), 2)))  # min/max longitude of the interpolation area
y.range <- as.numeric(c(round(min(data3$Y), 2), round(max(data3$Y), 2)))  # min/max latitude of the interpolation area

# GRID
# from the range, exapnd the coordinates to make a regular grid
grd <- expand.grid(x = seq(from = x.range[1], to = x.range[2], by = 0.001), y = seq(from = y.range[1], to = y.range[2], by = 0.001))
library(raster)
library(rgdal)
library(rgeos)
library(ggplot2)
library(dplyr)

### Q1: Delhi only

### Get data
india <- getData("GADM", country = "India", level = 1)

### Choose Delhi
delhi <- subset(india, NAME_1 == "NCT of Delhi")

### SPDF to DF
delhiline <- fortify(delhi)

grid <- makegrid(delhi, cellsize = 0.001) # cellsize in map units!
grid <- SpatialPoints(grid, proj4string = CRS(proj4string(delhi)))

grid <- grid[delhi, ]
grid <- as.data.frame(grid)

bestpar_matrix <- matrix(c(650, 1050, 0.15, 1000, 7100, 0.1, 250, 1350, 0.208, 1300, 3100, 0.141, 5500, 7700, 0.2, 1000, 3300, 0.173, 788.37, 4200, 0.224, 2550, 8400, 0.075), nrow = 8, ncol = 3, byrow = TRUE)

# KRIGING

dist <- pairdist(data3$X, data3$Y, na.ignore = T)

# min <- +Inf
# max <- -Inf

for(i in 1:8){
  j <- i+2
  a <- bestpar_matrix[i, 3]
  c0 <- bestpar_matrix[i, 1]
  c1 <- bestpar_matrix[i, 2] - c0
  

  covariance1 <- ifelse(dist > a, 0, c1*(1 -(1.5*abs(dist)/a)+(0.5*(abs(dist)/a)^3)))

  covariance1 <- cbind(covariance1, rep(1, nrow(covariance1)))
  covariance1 <- rbind(covariance1, rep(1, ncol(covariance1)))
  covariance1[nrow(covariance1), ncol(covariance1)] <- 0

  covinverse <- inv(covariance1)       #finds inverse of the matrix

  #dist23 <- proxy::dist(data3[1:2], grid[1:2])
  dist2 <- fields::rdist(data3[1:2], grid[1:2])

  covariance2 <- ifelse(dist2 > a, 0, c1*(1 -(1.5*abs(dist2)/a)+(0.5*(abs(dist2)/a)^3)))
  # covariance2 <- rbind(covariance2, 1)

  data4 <- t(rbind(data3[j], 0))
  data4[is.na(data4)] <- 0
  d <- rbind(covariance2, 1)
  w <- covinverse %*% d
  predicted_data_2 <- data4 %*% w
  predicted_data_2 <- sapply(predicted_data_2, function(x) ifelse(x < 0, 0, round(x, 2)))
  #print(w)

  kriging.output <- data.frame(as.vector(grid$x1), as.vector(grid$x2), predicted_data_2)
  colnames(kriging.output) <- c("grd.x", "grd.y", "PM2.5")

  # IDW

  # predicted_data <- cbind(predict_data[,c(1,2,j)], "Predicted_PM2.5_Kriging" = 0)
  # predicted_data["Predicted_PM2.5_IDW"] <- round(phylin::idw(data3[,j], data3[,c(1,2)], predict_data[,c(1,2)]), 2)

  interpolated <- phylin::idw(data3[,j], data3[,c(1,2)], grid)
  idw.output <- data.frame(as.vector(grid$x1), as.vector(grid$x2), as.vector(interpolated))
  colnames(idw.output) <- c("grd.x", "grd.y", "PM2.5")
  
  # if(min > min(kriging.output$PM2.5))
  #   min <- min(kriging.output$PM2.5)
  # if(min > min(idw.output$PM2.5))
  #   min <- min(idw.output$PM2.5)
  # 
  # if(max < max(kriging.output$PM2.5))
  #   max <- max(kriging.output$PM2.5)
  # if(max < max(idw.output$PM2.5))
  #   max <- max(idw.output$PM2.5)

  min_value_row <- c(kriging.output[nrow(kriging.output),1], kriging.output[nrow(kriging.output),2]+0.001, min)
  max_value_row <- c(kriging.output[nrow(kriging.output),1]+0.001, kriging.output[nrow(kriging.output),2], max)
  kriging.output <- rbind(kriging.output, min_value_row, max_value_row)

  min_value_row <- c(idw.output[nrow(idw.output),1], idw.output[nrow(idw.output),2]+0.001, min)
  max_value_row <- c(idw.output[nrow(idw.output),1]+0.001, idw.output[nrow(idw.output),2], max)
  idw.output <- rbind(idw.output, min_value_row, max_value_row)

  ti <- paste("Kriging")
  print(ggplot() +
    geom_tile(data = kriging.output, aes(x = grd.x, y = grd.y, fill = PM2.5)) +
    geom_point(data = data3, aes(x = X, y = Y), shape = 19, color = "black") +
    geom_point(data = predict_data, aes(x = X, y = Y), shape = 19, color = "black") +
    scale_fill_distiller(palette = "Spectral", direction = -1) +
    theme_bw() +

    labs(title = ti))
  
  ti <- paste("IDW")
  print(ggplot() +
    geom_tile(data = idw.output, aes(x = grd.x, y = grd.y, fill = PM2.5)) +
    geom_point(data = data3, aes(x = X, y = Y), shape = 19, color = "black") +
    geom_point(data = predict_data, aes(x = X, y = Y), shape = 19, color = "black") +
    scale_fill_distiller(palette = "Spectral", direction = -1) +
    theme_bw() +

    labs(title = ti))
}

# Palettes
# Diverging - BrBG, PiYG, PRGn, PuOr, RdBu, RdGy, RdYlBu, RdYlGn, Spectral
# Qualitative - Accent, Dark2, Paired, Pastel1, Pastel2, Set1, Set2, Set3
# Sequential - Blues, BuGn, BuPu, GnBu, Greens, Greys, Oranges, OrRd, PuBu, PuBuGn, PuRd, Purples, RdPu, Reds, YlGn, YlGnBu, YlOrBr, YlOrRd
```

<font size = 5>**There are many things which can be observed/inferred from the above variogram, in general:**<br>
<ol><li> The moment of inertia/semivariance is less for low distance, which is understandable as there will be many data points close to each other (for low distances).

<li> Theoretically, as distance increases, data points drift away from the line _x=y_ on an h-scatterplot so the cloud of data points will get fatter and MOI increases. This can be verified from this variogram. As the distance increases, the MOI also increases (roughly).</ol></font>

```{r}
# ERROR METRICS
library(ModelMetrics)
library(MLmetrics)
mae1 <- mae(predicted_data$PM2.5, predicted_data$Predicted_PM2.5_Kriging)
mae2 <- mae(predicted_data$PM2.5, predicted_data$Predicted_PM2.5_IDW)

my_RMSE_1 <- caret::RMSE(predicted_data$Predicted_PM2.5_Kriging, predicted_data$PM2.5)
my_RMSE_2 <- caret::RMSE(predicted_data$Predicted_PM2.5_IDW, predicted_data$PM2.5)

mape1 <- MAPE(predicted_data$Predicted_PM2.5_Kriging, predicted_data$PM2.5)
mape2 <- MAPE(predicted_data$Predicted_PM2.5_IDW, predicted_data$PM2.5)

df1 <- data.frame("Ordinary Kriging" = c(mae1, my_RMSE_1, mape1), "IDW Interpolation" = c(mae2, my_RMSE_2, mape2))
colnames(df1) <- c("Ordinary Kriging", "IDW Interpolation")
rownames(df1) <- c("Mean Absolute Error (MAE)", "Root Mean Square Error (RMSE)", "Mean Absolute Percentage Error (MAPE)")

df1 <- round(df1, 2)

df1[3,] <- lapply(df1[3,], function(x) paste(as.numeric(x)*100, "%", sep = ""))

```

```{r echo = FALSE, warning = FALSE}
library(knitr)
library(kableExtra)
options(knitr.table.format = "html")
kable(df1, align="c", "html") %>%
kable_styling("striped",full_width = F) %>%
row_spec(1, color = "black", background = "#ffe6e6") %>%
row_spec(2, color = "black", background = "#ffff99") %>%
row_spec(3, color = "black", background = "#ffe6e6") %>%
column_spec(1, bold = T, border_right = T) %>%
column_spec(2, border_right = T)

```
